---
title: "Quality Control of Respiration measurements"
format: html
editor: visual
---

```{r}
#| message: false
#| warning: false
#| include: false
knitr::opts_chunk$set(echo = T, eval = T, message = F, warning = F)
# load libraries
library(tidyverse)
library(here)
library(lubridate)
library(ggrepel)
library(broom)
library(readxl)

# Set ggplot theme
theme_set(theme_bw())
```

## 

This script provides a guide for how to perform QA/QC on respiration measurements

## User setup

**Instructions for SEI users**: 1) Download the "\*-Respiration.xlsx" spreadsheets to a folder called something like `data/licor/` or `data/picarro/` depending on the instrument that the measurements came from. All respiration spreadsheets should have names that start with a date in `YYYY-MM-DD` format, and end with `_Respiration.xlsx`

2)  Specify the location of spreadsheets and jar metadata as well. The `26SEI_JarLabelingandLoadingSpreadsheet_all_jars.tsv` combines both the Healy and Ice Cut samples and is generated by the utility script "SEI_metadata_reformat.R" (found in the `utilities` folder). To run the utility script you'll need the "26SEI_JarLabelingandLoadingSpreadsheet.xlsx" and "GasMeasurementTracking_SEI26(Flush 1).csv" tables from sharepoint

**Note: It is poor practice to place downloaded files in "Downloads", we recommend changing the output and input directories**

```{r user-defined-variables}
# the location you plots and tables generated by this code to be saved
fp_outputs <- "~/Downloads/SEI2026_respirationQC"

# location of sample metadata sheet
fp_metadata <- "~/Downloads/26SEI_JarLabelingandLoadingSpreadsheet_all_jars.tsv"

# location of sample respiration spreadsheets
fp_licor <- "~/Downloads/OneDrive_1_2-2-2026/LicorData_PreInc_SEI26/"

# location of sample respiration spreadsheets
fp_picarro <- "~/Downloads/OneDrive_1_2-2-2026/PicarroData_PreInc_SEI26/" # temporary; eventually move to examples

# location of the Ernakovich 2017 data
use_ernakovich_data <- TRUE # set to false if you don't want to generate plots with the Ernakovich data
ernakovich2017_fp <- here("examples/incubation/Ernakovich2017_data/Ernakovich_2017_data.xlsx")


```

```{r useful-constants-functions}

# Define Constants
GasConstant_R <- 0.0821 # L*atm*K^-1*mol^-1
PSI_to_atm <- 0.06804596 # psi to atm
PSI_to_hPa <- 68.9476 # psi to hpa
hPa_to_atm <- 0.0009869233 # hpa to atm 

# TEA levels
tea_levels = c("Carbon Dioxide", "Sulfate", "Iron", "Nitrate", "Oxygen", "None")
# Temperature levels
temp_levels = c("4", "10", "15", "20", "25", "None")
temp_colors = hcl.colors(length(temp_levels)-1, palette = "Berlin")
temp_colors = c(temp_colors, "grey") # adds the "None"
# Outlier function
isnt_out_tukey <- function(x, k = 1.5, na.rm = TRUE) {
  quar <- quantile(x, probs = c(0.25, 0.75), na.rm = na.rm)
  iqr <- diff(quar)
  
  (quar[1] - k * iqr <= x) & (x <= quar[2] + k * iqr)
}
```

```{r, make-output-dirs}
dir.create(fp_outputs)

# Figures output directory
fp_figout <- paste0(fp_outputs, "/figures/")
dir.create(fp_figout)

# tables output directory
fp_tabout <- paste0(fp_outputs, "/tables/")
dir.create(fp_tabout)

```

### Reading in the sample metadata and jar information

First we'll read in the sample metadata and jar information. This is important because it contains the pre-incubation and incubation start times.

```{r read-sample-data}
# Reading in sample metadata information
jar_info_raw <- read_tsv(file = fp_metadata) %>%
  mutate(across(JarNumber:`Specimen Cup Number`, ~ as.factor(.x))) %>%
  rename(AnoxicOxic =`Anoxic/Oxic?`) %>%
  mutate(TEA = factor(TEA, levels =tea_levels),
         `Incubation Temp` = factor( `Incubation Temp`, levels = temp_levels))

# Extract incubation start times
jar_incubation_time_info <- jar_info_raw %>%
  select(JarIDNum, contains("IncubationStart")) %>%
  mutate(PreIncubationStart = force_tz(PreIncubationStart, tz = "America/New_York"),
         IncubationStart = force_tz(IncubationStart, tz = "America/New_York")) # temporary until start date is determined added in

# Extract data relevant for QC of samples, plotting, and calculation of rates 
jar_info <- jar_info_raw %>%
  select(SampleID:Site, `Incubation Temp`, `Mass of Soil(g Lab Scale)`, Mass_of_PreInc_subsample_g, GSC_Perc, HeadspaceVolume)
```

If you set the options above to use the Ernakovich 2017 data, we'll read it in and prepare it for comparison.

```{r read-ernakovich2017}
if(use_ernakovich_data){
ernakovich_2017_CO2 <- readxl::read_excel(ernakovich2017_fp, 
                                      sheet = 1, 
                                      range = cell_limits(c(2,1), c(16,9)),
                                      na = c("nan", "#VALUE!", "#DIV/0!")) %>%
  select(`Day of Incubation`, contains("oxic")) %>%
  pivot_longer(-`Day of Incubation`, names_to = "IncubationTreatmentAndStat", values_to = "CO2_ugC_gdws_hr") %>%
  mutate(MeanSE = gsub("(^.{1,})_(.{1,}$)", "\\2", IncubationTreatmentAndStat),
         Treatment = gsub("(^.{1,})_(.{1,}$)", "\\1", IncubationTreatmentAndStat)) %>%
  select(-IncubationTreatmentAndStat) %>%
  pivot_wider(names_from = "MeanSE",
              values_from = "CO2_ugC_gdws_hr") %>%
  # Create equivalent columns
  mutate(ug_C_CO2_per_g_soil_per_hour = mean,
         ug_C_CO2_per_g_soil_per_hour_se = se,
         ug_C_CO2_per_g_soil_per_day = mean*24,
         ug_C_CO2_per_g_soil_per_day_se = se*24) %>%
  select(-mean, -se)

ernakovich_2017_CH4 <- readxl::read_excel(ernakovich2017_fp, 
                                      sheet = 1, 
                                      range = cell_limits(c(20,1), c(34,9)),
                                      na = c("nan", "#VALUE!", "#DIV/0!")) %>%
  select(`Day of Incubation`, contains("oxic")) %>%
  pivot_longer(-`Day of Incubation`, names_to = "IncubationTreatmentAndStat", values_to = "CH4_ugC_gdws_hr") %>%
  mutate(MeanSE = gsub("(^.{1,})_(.{1,}$)", "\\2", IncubationTreatmentAndStat),
         Treatment = gsub("(^.{1,})_(.{1,}$)", "\\1", IncubationTreatmentAndStat)) %>%
  select(-IncubationTreatmentAndStat) %>%
  pivot_wider(names_from = "MeanSE",
              values_from = "CH4_ugC_gdws_hr") %>%
    mutate(ug_C_CH4_per_g_soil_per_hour = mean,
         ug_C_CH4_per_g_soil_per_hour_se = se,
         ug_C_CH4_per_g_soil_per_day = mean*24,
         ug_C_CH4_per_g_soil_per_day_se = se*24) %>%
  select(-mean, -se)


ernakovich_2017 <- left_join(ernakovich_2017_CO2, ernakovich_2017_CH4, by = c("Day of Incubation", "Treatment"))  
}

```

Below we create a function for reading in respiration data and doing basic reformatting that allows the picarro and licor respiration measurements to be combined. These include: 1. Converting picarro readings from 12-C and 13-C into total C 2. Making sure all the columns have the right formats in R 3. Ensuring the dates and times are correctly formatted, and in the right time zone 4. Converting licor and picarro information into each other with a scaling factor. *Does not do this currently - but will eventually*

```{r, read-resp-data}
read_respiration_files <- function(resp_file_path_licor = NULL,
                                   resp_file_path_picarro = NULL, 
                                   sample_id_column = "Jar Number",
                                   incubation_start=jar_incubation_time_info,
                                   file_pattern = ".xlsx") {
#' Function: read_respiration_files()
#' Takes a folder containing either Picarro or Licor-style respiration templates
#' and reads them in, does baisc cleaning functions on them. 
# @param resp_file_path_licor a directory path where identically formatted respiration spreadsheets for the licor are housed. If you do not have data from the instrument, leave argument blank.
# @param resp_file_path_picarro a directory path where identically formated respiration spreadsheets for the picarro are housed. If you do not have data from the instrument, leave argument blank.
# @param sample_id_column defaults to "Jar Number", change only if you have changed the sample id column name. A character.
# @param incubation_start A dataframe that includes the sample ID column and the time of the incubation start. 
# @return a list of data frames containing the licor data, the picarro data, and a data frame that combines both.
# @examples tbd
  
  # Dependencies
  library(tidyverse)
  library(lubridate)
  library(readxl)
  
  # Constants
  PSI_to_atm <- 0.06804596 # for converting headspace pressure to moles of gas
  
  # Return list
  return_list <- NA
  
  # Testing
  # resp_file_path_licor = "~/Downloads/licor_data"
  # resp_file_path_picarro = "~/Downloads/OneDrive_1_1-20-2026"
  # incubation_start <- "2026-01-09 12:00:00" 
  # file_pattern = ".xlsx"
  # Check arguments for validity and warn user about states  
  stopifnot((!is.null(resp_file_path_licor) | !is.null(resp_file_path_picarro)))
  
  if(!is_null(resp_file_path_licor)) {
    # print("here") # Testing
    Resp_files <- list.files(here(resp_file_path_licor),
                             pattern = file_pattern, full.names = T)
    
    col_types_licor <- c(rep("text", 2), #SampleType:JarNumber
                         rep("numeric", 5), #RoomTemperature:ConcentrationOfStandard
                         "date", # Time last flushed
                         rep("text", 3), # Comments, id, si.no
                         "date", # timestamp
                         rep("numeric", 4), # mean-pre:Est Conc
                         "text", # Remarks
                         rep("numeric", 10), #:effective volume of loop:True amount of CO2
                       "guess","text", "numeric", "numeric") # standard curve calcs
    
    resp_raw_licor <- map_dfr(Resp_files, 
                               ~readxl::read_excel(.x,sheet = 1,
                                                   col_types = col_types_licor,
                                                   range = cell_limits(c(1,1), c(NA,31)),
                                                   na = c("nan", "#VALUE!", "#DIV/0!")) %>%
                                rename_with(.fn = ~ gsub("\\(psi\\)", "\\(psig\\)", .)))
    resp_raw_licor <- resp_raw_licor %>%
      filter(!is.na(SampleType)) %>%
      # Modify jar names to get jar id number
      # When a pattern of word-word-number is detected, extract the number
      # When zero hyphens are encountered use the number as written
      # when anything else is encountered use the number as written
      mutate(JarIDNum = case_when(grepl("(.{1,})-(.{1,})-(\\d{1,3})", 
                                        `Jar Number`) ~ str_extract(`Jar Number`,
                                                                    "(.{1,})-(.{1,})-(\\d{1,3})",
                                                                    group = 3),
                                  grepl("(C.{1,})-(.{1,})-(.{1,})", 
                                        `Jar Number`) ~str_extract(`Jar Number`,
                                                                    "(C.{1,})-(.{1,})-(.{1,})",
                                                                    group = 1),
                                  str_count(`Jar Number`, "-") == 0 ~ `Jar Number`,
                                  .default = `Jar Number`)) %>%
      select(SampleType,JarIDNum, `Jar Number`:`True amount of CO2 (uMol)`) %>%
      # Add in measurement instrument and fix the timezone of time columns
      mutate(MeasurementInstrument = "Licor",
             `Time Incubation Jar was Last Flushed (Leave blank for standard curve)` = force_tz(`Time Incubation Jar was Last Flushed (Leave blank for standard curve)`, tz = "America/New_York"), 
             Timestamp = force_tz(Timestamp, tz = "America/New_York")) # need to check that this is correct!
    
    # Change names to match variables between licor and picarro
    resp_raw_licor <- resp_raw_licor %>%
      mutate(`TotalCO2 (ppm)` = `True amount of CO2 (uMol)`/ `Moles of Air in sample`) %>%
      rename(`Concentration of Standard Curve CO2 Gas (ppm)` = `Concentration of Standard Curve Gas (ppm)`,
             `TotalCO2 (uMol)` = `True amount of CO2 (uMol)`) %>%
      # Calculate Jar pressure
      mutate(AbsoluteJarPressure_hPa = `RoomPressure (hPa)` + `Jar pressure before sampling (psig)`*PSI_to_hPa,
             AbsoluteJarPressure_atm = AbsoluteJarPressure_hPa * hPa_to_atm) %>%
      # calcualte durations:
      left_join(incubation_start, by = c("JarIDNum")) %>%
      mutate(Date = as_date(Timestamp),
             IncubationDay = Date - as_date(IncubationStart),
             IncubationTime = Timestamp - as_datetime(IncubationStart, tz = "America/New_York"),
             PreIncubationDay = Date - as_date(PreIncubationStart),
             PreIncubationTime = Timestamp - as_datetime(PreIncubationStart, tz = "America/New_York"),
             TimeSinceZerod = Timestamp - `Time Incubation Jar was Last Flushed (Leave blank for standard curve)`)
  }
  
  if(exists("resp_file_path_picarro")) {
    Resp_files <- list.files(here(resp_file_path_picarro),
                             pattern = file_pattern, full.names = T)
    
    col_types_picarro <- c(rep("text", 2), #sampletype:jar number
                       rep("numeric", 16), #RoomTemperature:EffectiveIsotopicRatio
                       "date", #date last flushed
                       "date", #time last flushed
                       rep("text", 1), # comments
                       "date", # Date
                       "date", # time of start of measurements
                       rep("numeric", 8), #Mean of pre-12CO2:MeanH2O%post 
                       "text", # Remarks 
                       rep("guess", 32), # effective volume of loop:TrueDelta permil
                       rep("guess", 12)) # regression calculations
    # readxl::read_excel(Resp_files[2], sheet = 1, skip = 1,
    #                                                col_types = col_types_picarro,
    #                                                na = c("nan", "#VALUE!", "#DIV/0!", "#REF!"))
    
    # Read in data
    resp_raw_picarro <- map_dfr(Resp_files, 
                               ~readxl::read_excel(., sheet = 1,
                                                   col_types = col_types_picarro,
                                                   range = cell_limits(c(2,1), c(NA,76)),
                                                   na = c("nan", "#VALUE!", "#DIV/0!", "#REF!", "NA",
                                                          "Invalid")) %>%
                                rename_with(.fn = ~ gsub("\\(psi\\)", "\\(psig\\)", .))) %>% 
      filter(!is.na(SampleType)) %>%
      # Modify jar names to get jar id number
      # When a pattern of word-word-number is detected, extract the number
      # When zero hyphens are encountered use the number as written
      # when anything else is encountered use the number as written
      mutate(JarIDNum = case_when(grepl("(.{1,})-(.{1,})-(\\d{1,3})", 
                                        `Jar Number`) ~ str_extract(`Jar Number`,
                                                                    "(.{1,})-(.{1,})-(\\d{1,3})",
                                                                    group = 3),
                                  grepl("(C.{1,})-(.{1,})-(.{1,})", 
                                        `Jar Number`) ~str_extract(`Jar Number`,
                                                                    "(C.{1,})-(.{1,})-(.{1,})",
                                                                    group = 1),
                                  str_count(`Jar Number`, "-") == 0 ~ `Jar Number`,
                                  .default = `Jar Number`)) %>%
      select(SampleType, JarIDNum, `Jar Number`:`True delta iCO2 (permil)`) %>%
      mutate(MeasurementInstrument = "picarro")
    
    # Fix timestamps for measurement and flush time
    date(resp_raw_picarro$`Time of start of measurement`) <- resp_raw_picarro$Date 
    resp_raw_picarro$`Time of start of measurement` <- force_tz(resp_raw_picarro$`Time of start of measurement`, "America/New_York") 
    
    date(resp_raw_picarro$`Time Incubation Jar was Last Flushed (Leave blank for standard curve)`) <- resp_raw_picarro$`Date Incubation Jar was Last Flushed (Leave blank for standard curve)` 
    resp_raw_picarro$`Time Incubation Jar was Last Flushed (Leave blank for standard curve)` <- force_tz(resp_raw_picarro$`Time Incubation Jar was Last Flushed (Leave blank for standard curve)`, "America/New_York") 
    
    # Create columns for total CH4 and CO2
    resp_raw_picarro <- resp_raw_picarro %>% 
      mutate(`atom%12CO2 based on delta` = 1 - `Uncalibrated apparent atom% 13CO2 based on delta value`,
             `atom%12CH4 based on delta` = 1 - `Uncalibrated apparent atom% 13CH4 based on delta value`,
             `TotalCO2 (ppm)` = `Uncalibrated apparent 12CO2 concentration (ppm)` / `atom%12CO2 based on delta`,
             `TotalCH4 (ppm)` = `Uncalibrated apparent 12CH4 concentration (ppm)` / `atom%12CH4 based on delta`,
             `TotalCO2 (uMol)` = `TotalCO2 (ppm)` * `Moles of Air in sample`,
             `TotalCH4 (uMol)` = `TotalCH4 (ppm)` * `Moles of Air in sample`) %>%
      # Create one timestamp
      rename(Timestamp = `Time of start of measurement`) %>%
      # convert jar pressure to atm:
      mutate(AbsoluteJarPressure_hPa = `RoomPressure (hPa)` + `Jar pressure before sampling (psig)`*PSI_to_hPa,
             AbsoluteJarPressure_atm = AbsoluteJarPressure_hPa * hPa_to_atm) %>%
      # calculate durations
      left_join(incubation_start, by = c("JarIDNum")) %>%
      mutate(Date = as_date(Timestamp),
             IncubationDay = Date - as_date(IncubationStart),
             IncubationTime = Timestamp - as_datetime(IncubationStart, tz = "America/New_York"),
             PreIncubationDay = Date - as_date(PreIncubationStart),
             PreIncubationTime = Timestamp - as_datetime(PreIncubationStart, tz = "America/New_York"),
             TimeSinceZerod = Timestamp - `Time Incubation Jar was Last Flushed (Leave blank for standard curve)`)
  }
  
  
  # Next steps:
  # Determine which of the resp_raw data is present (use exists(resp_raw_picarro))
  # combine data together into one dataframe
  # Create the incubation day and time columns, use either the user-specified beginning datetime or the minimum date
  
  if(exists("resp_raw_licor") & exists("resp_raw_picarro")) {
    resp_raw <- bind_rows(resp_raw_licor, resp_raw_picarro)
    return_list <- list(licor_resp_raw = resp_raw_licor,
                        picarro_resp_raw = resp_raw_picarro)
  } else if(exists("resp_raw_licor")) {
    resp_raw <- resp_raw_licor
    return_list <- list(licor_resp_raw = resp_raw_licor,
                    picarro_resp_raw = NA)
  } else if(exists("resp_raw_picarro")) {
      resp_raw <- resp_raw_picarro
      return_list <- list(licor_resp_raw = NA,
                    picarro_resp_raw = resp_raw_picarro)
  }
  
  
  # Calculate durations and jar pressures
    resp_raw <- resp_raw %>%  
      mutate(SampleType = factor(SampleType),
             `Jar Number` = factor(as.character(`Jar Number`)),
             `JarIDNum` = factor(as.character(`JarIDNum`)))
  
  # Return statement
  return_list$resp_raw <- resp_raw
  return(return_list)  
}

```

Run the function to load the input.

```{r}
resp_dat_raw <- read_respiration_files(
  resp_file_path_licor = fp_licor,
  resp_file_path_picarro = fp_picarro,
  incubation_start = jar_incubation_time_info,
  file_pattern = ".xlsx")

```

**STOP:** If there are any warnings or errors, check the spreadsheets you are reading in before proceeding.

### Calculate Respirations

Next we need to calculate the respiration rates. In brief this means we need to:

1.  Correct the amount of gas measured for the volume of air in the headspace (which can vary between jars - particularly if you use different brands)
    -   calculate convert the jar pressure from PSIG (the guage air pressure) to atm
2.  Calculate the dry soil weight in each jar (based on the gravimetric water content or gravimetric soil content (`GSC_perc`) )
3.  Convert the total gas produced in micromoles to micrograms (to align with conventional units)
4.  Divide the amount of gas produced by the hours and days since the jars have been flushed to produce and hourly and daily rate

```{r, respiration-calculations}

samples_resp <- resp_dat_raw$resp_raw %>%
  filter(SampleType == "Sample") %>%
  # make samples within precision of instrument zero
  mutate(`TotalCH4 (ppm)` = case_when(`Difference in 12CH4 quality flag` == "DifferenceBelowPrecision" ~ 0,
                                      .default = `TotalCH4 (ppm)`)) %>%
  mutate(`TotalCH4 (uMol)` = case_when(`Difference in 12CH4 quality flag` == "DifferenceBelowPrecision" ~ 0,
                                      .default = `TotalCH4 (uMol)`)) %>%
  filter(if_any(contains("TotalCH4"), ~(is.na(.x) | .x >=0))) %>%
  # Convert into a per-gram basis, based on Jar number
  left_join(jar_info, by = c("JarIDNum")) %>%
  mutate(jar_headspace_mL = HeadspaceVolume, # need to correct for specimen cup volume
         Mass_after_preinc_subsample = `Mass of Soil(g Lab Scale)` - Mass_of_PreInc_subsample_g,
         Mass_of_soil_at_time_of_gas_measurement = case_when(is.na(IncubationDay) ~ `Mass of Soil(g Lab Scale)`,
                                                             IncubationDay < days(0) ~ `Mass of Soil(g Lab Scale)`,
                                                             IncubationDay >= days(0) ~ Mass_after_preinc_subsample, 
                                                             .default = NA_real_)) %>%
  mutate(soil_weight_dry = Mass_of_soil_at_time_of_gas_measurement * (GSC_Perc/100),
         MolAir_headspace = (1/1000)*(jar_headspace_mL * AbsoluteJarPressure_atm)/(GasConstant_R * `Room Temperature (K)`),
         umol_CO2_in_headspace = jar_headspace_mL * (`TotalCO2 (uMol)`)/`Volume of sample or standard (mL of gas measured)`,
         umol_CH4_in_headspace = jar_headspace_mL * (`TotalCH4 (uMol)`)/`Volume of sample or standard (mL of gas measured)`,
         umol_CO2_per_molAir = umol_CO2_in_headspace/MolAir_headspace,
         umol_CH4_per_molAir = umol_CH4_in_headspace/MolAir_headspace,
         ug_C_CO2_in_headspace = jar_headspace_mL * (`TotalCO2 (uMol)` * 12.01)/`Volume of sample or standard (mL of gas measured)`,
         ug_C_CH4_in_headspace = jar_headspace_mL * (`TotalCH4 (uMol)` * 12.01)/`Volume of sample or standard (mL of gas measured)`,
         ug_C_CO2_per_g_soil = ug_C_CO2_in_headspace/soil_weight_dry,
         ug_C_CH4_per_g_soil = ug_C_CH4_in_headspace/soil_weight_dry) %>%
  # make it a rate just using temporary numbers here
  mutate(ug_C_CO2_per_g_soil_per_hour = ug_C_CO2_per_g_soil/as.numeric(TimeSinceZerod, unit = "hours"),
         ug_C_CO2_per_g_soil_per_day = ug_C_CO2_per_g_soil/(as.numeric(TimeSinceZerod, unit = "days"))) %>%
  mutate(ug_C_CH4_per_g_soil_per_hour = ug_C_CH4_per_g_soil/as.numeric(TimeSinceZerod, unit = "hours"),
         ug_C_CH4_per_g_soil_per_day = ug_C_CH4_per_g_soil/(as.numeric(TimeSinceZerod, unit = "days"))) %>%
    mutate(PercentCH4 = `ug_C_CH4_per_g_soil`/(`ug_C_CH4_per_g_soil` + `ug_C_CO2_per_g_soil`))


```

Placeholder for future checks where we can compare internal standard values to expected values.

```{r, internal-standards}
standard_readings <- resp_dat_raw$resp_raw %>%
  filter(SampleType == "Internal Standard") %>%
  select(Timestamp, `Jar Number`, `TotalCO2 (ppm)`, `TotalCH4 (ppm)`) %>%
  mutate(DistanceFromStandard = case_when(grepl("ethane", `Jar Number`) ~ 1000-`TotalCH4 (ppm)`,
                                             grepl("1022", `Jar Number`) ~ 1022-`TotalCO2 (ppm)`,
                                             grepl("CO2", `Jar Number`) ~ 1022-`TotalCO2 (ppm)`,
                                             grepl("Tank", `Jar Number`)~ 1022-`TotalCO2 (ppm)`,
                                             .default = NA_real_))

standard_readings %>%
  filter(`TotalCO2 (ppm)` < 5000) %>%
  ggplot(aes(x = Timestamp, y = DistanceFromStandard)) +
  geom_point() +
  geom_text_repel(aes(label = `Jar Number`)) +
  geom_hline(aes(yintercept = 0)) +
  scale_x_date(date_breaks = "20 days")

acceptable_tolerance <- 0.05 # 5%
internal_standard_tolerance_df <- tibble(
  Standard = c("Methane 1000", "CO2 1022"),
  GasMeasured = c("TotalCH4 (ppm)", "TotalCO2 (ppm)"),
  `Concentration (PPM)` = c(1000, 1022),
  upper_limit = c(1000*(1 +acceptable_tolerance), 1022*(1 +acceptable_tolerance)),
  lower_limit = c(1000*(1 -acceptable_tolerance), 1022*(1 - acceptable_tolerance)),
  Timestamp = NA
)
standard_readings %>%
    filter(`TotalCO2 (ppm)` < 5000) %>%
  pivot_longer(contains("ppm"), names_to = "GasMeasured", values_to = "Concentration (PPM)") %>%
  ggplot(aes(x = Timestamp, y = `Concentration (PPM)`)) +
  geom_point() +
  geom_text_repel(aes(label = `Jar Number`)) +
  geom_hline(data = internal_standard_tolerance_df, 
             aes(yintercept = `Concentration (PPM)`)) +
  geom_hline(data = internal_standard_tolerance_df, 
             aes(yintercept = lower_limit), color = "firebrick", linetype = "dashed") +
  geom_hline(data = internal_standard_tolerance_df, 
             aes(yintercept = upper_limit), color = "firebrick", linetype = "dashed") +
  facet_wrap(~GasMeasured)# +
  #scale_x_date(date_breaks = "20 days")


standard_readings %>%
  filter(`TotalCO2 (ppm)` < 5000) %>%
  ggplot(aes(x = Timestamp, y = `TotalCO2 (ppm)`)) +
  geom_point() +
  geom_text_repel(aes(label = `Jar Number`)) +
  geom_hline(aes(yintercept = 1022)) +
  scale_x_date(date_breaks = "20 days")
```

# Basic checks:

1)  Any samples NA? figure out why before moving on:

```{r}
rate_cols <- c("ug_C_CO2_per_g_soil_per_hour", "ug_C_CO2_per_g_soil_per_day") 

picarro_rate_cols <- c("ug_C_CH4_per_g_soil_per_hour","ug_C_CH4_per_g_soil_per_day") 

# Check co2 measurements
samples_resp %>%
  select(SampleType, JarIDNum, MeasurementInstrument, Timestamp, 
         all_of(rate_cols), everything()) %>%
  filter(if_any(all_of(rate_cols), ~is.na(.x))) %>% View()

# Check picarro-only measurements
samples_resp %>%
  filter(MeasurementInstrument == "picarro") %>%
  select(SampleType, JarIDNum, MeasurementInstrument, Timestamp, 
         all_of(picarro_rate_cols), everything()) %>%
  filter(if_any(all_of(picarro_rate_cols), ~is.na(.x))) %>% View()


# Check for measurements below instrument precision
samples_resp %>%
  filter(MeasurementInstrument == "picarro") %>%
  select(SampleType, JarIDNum, MeasurementInstrument, Timestamp, 
         all_of(picarro_rate_cols), everything()) %>%
  filter(if_any(all_of(contains("quality flag")), ~ grepl("Below", .x))) %>% View()

```

Common issues:

-   Flush date and time are not entered correctly
    -   If so, leave a comment in the spreadsheet tagging the person responsible for the measurement.
-   temperature, pressure, or other necessary variable is missing.
-   soil weights not entered
-   Typo in sample name or number
-   Measurement is impossible (e.g. negative ppm); if so, update data spreasheet on sharepoint with note in comments
-   equations need to be copied into cells in spreasheet

If necessary, find the date of the timestamp for the measurement and check the corresponding excel sheet and lab notebook pages. Message the person who took the measurement to verify any uncertainties.

**Only make changes to the raw spreadsheets if you are certain you know the correct value. And always note your changes in a comment column, with your name or initials and the date.**

2)  Check for duplicate observations on the same day. Make sure that all are appropriate

-   appropriate duplicates include jars measured on both picarro and licor.

```{r}
samples_resp %>%
  select(`JarIDNum`, Date, MeasurementInstrument,
         ug_C_CO2_in_headspace, ug_C_CO2_per_g_soil_per_hour, 
         ug_C_CO2_per_g_soil_per_day) %>%
  dplyr::summarise(n = dplyr::n(), .by= c(`JarIDNum`, Date,
  MeasurementInstrument)) %>%
  dplyr::filter(n > 1L)

# if the above code detects duplicates, choose them to filter; ask 
# the person who ran the samples which duplicate is the preferred measurement
duplicates_to_filter <- c("377", "45", "177")

```

3)  Check that the jars that were supposed to be measured were measured. \[TBD until checklist spreadsheets exist\]

4)  Histogram for any obvious outliers. If something is very large or very small, check the raw data to be sure it isn't a typo.

```{r}
samples_resp %>%
  select(`Jar Number`, Date, all_of(rate_cols)) %>%
  pivot_longer(all_of(rate_cols), names_to = "Rates", values_to = "Rate Value") %>%
  ggplot(aes(x = Date, y = `Rate Value`)) +
  geom_boxplot(aes(group = Date), outliers = F) +
  geom_point(position = "jitter") +
  geom_text_repel(aes(label = `Jar Number`)) +
  facet_wrap(~Rates, scales = "free_y")

```

5)  Comparison between instruments

-   we want to make a conversion factor between each of the two instruments, but this also can serve as a good check.
-   if any samples are wildly off the 1:1 line, go look into the original data sheets and verify it is not an issue with the original data.

```{r}
list_to_filter_out <- c(duplicates_to_filter, 
                        "384" # filtering per discussion with Maggie D; outside range of standard curve
                        )

licor_picarro_comparison <- samples_resp %>%
   filter_out(JarIDNum %in% list_to_filter_out) %>%
   group_by(JarIDNum, Date) %>%
   add_tally(name = "num_obs") %>%
   ungroup() %>%
   filter(num_obs > 1) %>% arrange(Date, JarIDNum) %>% 
  # Filter only jars that are in list of jars
  filter(JarIDNum %in% jar_info$JarIDNum) %>%
  select(JarIDNum, Date, MeasurementInstrument,
         ug_C_CO2_in_headspace, ug_C_CO2_per_g_soil_per_hour, 
         ug_C_CO2_per_g_soil_per_day) %>%
  pivot_wider(names_from = MeasurementInstrument, 
              names_glue = "{MeasurementInstrument}_{.value}",
              values_from = c(ug_C_CO2_in_headspace,
                              ug_C_CO2_per_g_soil_per_hour, 
                              ug_C_CO2_per_g_soil_per_day)) %>%
  mutate(Date = as.factor(Date))

headspace_CO2_comp <- lm(picarro_ug_C_CO2_in_headspace ~ Licor_ug_C_CO2_in_headspace*Date, 
                         data = licor_picarro_comparison)
summary(headspace_CO2_comp)

headspace_CO2_comp <- lm(picarro_ug_C_CO2_in_headspace ~ Licor_ug_C_CO2_in_headspace, 
                         data = licor_picarro_comparison)
summary(headspace_CO2_comp)

hrly_CO2_comp <- lm(picarro_ug_C_CO2_per_g_soil_per_hour ~ Licor_ug_C_CO2_per_g_soil_per_hour*Date, 
                    data = licor_picarro_comparison)
summary(hrly_CO2_comp)

ggplot(licor_picarro_comparison, 
       aes(x = Licor_ug_C_CO2_in_headspace, y = picarro_ug_C_CO2_in_headspace)) +
  geom_point(aes(color = Date)) + 
  geom_text_repel(aes(label = `JarIDNum`)) +
  geom_abline(slope = 1,
              intercept = 0, color = "firebrick", linetype = "dashed") +
  geom_abline(slope = headspace_CO2_comp$coefficients[[2]],
              intercept = headspace_CO2_comp$coefficients[[1]]) +
  annotate(geom = "text", x  = -Inf, y = Inf, hjust = -0.10, vjust = 1.1,
           label = paste("y = ", round(headspace_CO2_comp$coefficients[[2]], digits = 2),"x + ",
                        round(headspace_CO2_comp$coefficients[[1]], digits = 2), "\n",
                        "R2 = ", round(summary(headspace_CO2_comp)$r.squared, digits = 4))
           ) +
  annotate(geom = "text", x  = 440, y = 450, hjust = 1,
           label = paste("1:1 line"), color = "firebrick")

ggplot(licor_picarro_comparison, 
       aes(x = Licor_ug_C_CO2_per_g_soil_per_hour, y = picarro_ug_C_CO2_per_g_soil_per_hour)) +
  geom_point(aes(color = as.factor(Date))) + 
  geom_text_repel(aes(label = `JarIDNum`)) +
  geom_abline(slope = hrly_CO2_comp$coefficients[[2]],
              intercept = hrly_CO2_comp$coefficients[[1]]) +
  geom_abline(slope = 1,
              intercept = 0, color = "firebrick", linetype = "dashed") +
  annotate(geom = "text", x  = -Inf, y = Inf, hjust = -0.10, vjust = 1.1,
           label = paste("y = ", round(hrly_CO2_comp$coefficients[[2]], digits = 2),"x + ",
                        round(hrly_CO2_comp$coefficients[[1]], digits = 2), "\n",
                        "R2 = ", round(summary(hrly_CO2_comp)$r.squared, digits = 4))
           )


licor_picarro_models <- licor_picarro_comparison %>%
  group_by(Date) %>%
  nest() %>%
  mutate(lm.model = map(data, ~lm(picarro_ug_C_CO2_in_headspace ~ Licor_ug_C_CO2_in_headspace, data = .)),
          lm.tidied = map(lm.model, tidy),
          lm.glanced = map(lm.model, glance),
          summaried = map(lm.model, summary),
          overall_p.value = map_dbl(lm.glanced, "p.value"),
          r.squared = map_dbl(lm.glanced, "r.squared")) %>%
  unnest("lm.tidied") %>%
  pivot_wider(names_from = "term", values_from = c(estimate, std.error, statistic, p.value))

#simplified - only slope intercept and date
licor_picarro_models_simplified <- licor_picarro_models %>%
  select(Date, `estimate_(Intercept)`, estimate_Licor_ug_C_CO2_in_headspace, r.squared) %>%
  rename(GasMeasurementDate = Date,
         LicorPicarroIntercept = `estimate_(Intercept)`,
         LicorPicarroSlope = estimate_Licor_ug_C_CO2_in_headspace) %>%
  mutate(GasMeasurementDate = as_date(GasMeasurementDate)) %>%
  ungroup()


ggplot(licor_picarro_comparison %>% mutate(Date = as.factor(Date)), 
       aes(x = Licor_ug_C_CO2_per_g_soil_per_hour, 
           y = picarro_ug_C_CO2_per_g_soil_per_hour, group = Date)) +
  geom_point(aes(color = Date)) + 
  geom_text_repel(aes(label = `JarIDNum`)) +
  facet_wrap(~Date) +
  geom_text(data = licor_picarro_models_simplified %>% mutate(Date = as.factor(GasMeasurementDate)), 
            x  = -Inf, y = Inf, hjust = -0.10, vjust = 1.1, 
            aes(label = paste("y = ", round(LicorPicarroSlope, digits = 2),"x + ",
                        round(LicorPicarroIntercept, digits = 2), "\n",
                        "R2 = ", round(r.squared, digits = 4)))
           ) +
  geom_abline(data = licor_picarro_models_simplified %>% 
                mutate(Date = as.factor(GasMeasurementDate),
                       Licor_ug_C_CO2_per_g_soil_per_hour = 0,
                       picarro_ug_C_CO2_per_g_soil_per_hour = 0),
              aes(slope = LicorPicarroSlope,
              intercept = LicorPicarroIntercept), inherit.aes = T)


```

Now that the relationship between licor and picarro is known for each measurement day, we can use it to re-calculate the licor output.

```{r}
samples_resp_instrument_corrected <- samples_resp %>%
  left_join(licor_picarro_models_simplified, by = c("Date" = "GasMeasurementDate")) %>%
  # for samples with no equivalent day licor readings, use the average slope of all days
  mutate(LicorPicarroIntercept = ifelse(!(Date %in% licor_picarro_models_simplified$GasMeasurementDate), headspace_CO2_comp$coefficients[[1]],
                                        LicorPicarroIntercept),
         LicorPicarroSlope = ifelse(!(Date %in% licor_picarro_models_simplified$GasMeasurementDate), headspace_CO2_comp$coefficients[[2]],
                                        LicorPicarroSlope)) %>%
  # Create correction columns:
  mutate( ug_C_CO2_in_headspace_inst_corrected = ifelse(MeasurementInstrument == "Licor", 
                                                        ug_C_CO2_in_headspace*LicorPicarroSlope + LicorPicarroIntercept, 
                                                        ug_C_CO2_in_headspace),
          ug_C_CO2_per_g_soil_inst_corrected = ug_C_CO2_in_headspace_inst_corrected/soil_weight_dry,
          ug_C_CO2_per_g_soil_per_hour_inst_corrected = ug_C_CO2_per_g_soil_inst_corrected/as.numeric(TimeSinceZerod, unit = "hours"),
         ug_C_CO2_per_g_soil_per_day_inst_corrected = ug_C_CO2_per_g_soil_inst_corrected/(as.numeric(TimeSinceZerod, unit = "days"))) %>%
  select(`SampleID`,JarIDNum,`Jar Number`, Date, contains("ug_C_CO2"), Site, 
         `Permafrost or Coalescence?`,
         PreIncubationDay, LicorPicarroIntercept, LicorPicarroSlope)

samples_resp_instrument_corrected %>%
  ungroup() %>%
  #filter(!grepl("Methane", `Jar Number`)) %>%
  separate(SampleID, into = c("SiteAbbr", "MeasureType", "JarNum"), 
           sep = "-", remove = F) %>%
  mutate(SiteAbbrPermCoal = paste(SiteAbbr, `Permafrost or Coalescence?`)) %>%
  filter(!is.na(SiteAbbrPermCoal)) %>%
  ggplot(aes(y = ug_C_CO2_per_g_soil_per_hour_inst_corrected, 
             x = PreIncubationDay, color = SiteAbbrPermCoal)) +
    geom_point(aes_string(shape = "Site"), size = 3) +
    geom_line(aes(group = SiteAbbrPermCoal))
# 
# samples_resp_instrument_corrected %>%
#   ggplot(aes(y = ug_C_CO2_per_g_soil_per_day_inst_corrected, x =ug_C_CO2_per_g_soil_per_day)) +
#   geom_point(aes(color = MeasurementInstrument))

```

# Plot respiration

Create a function for plotting respiration

```{r}
plot_resp <- function(resp_dat = samples_resp,
                      yval = "ug_C_CH4_per_g_soil",
                      xval = "PreIncubationDay",
                      summary_plot = T,
                      jar_labels = F,
                      all_points = T,
                      save = T,
                      save_location = fp_figout,
                      GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal")) {
  #args_list <- list(...)
  #print(args_list)
  #print(GroupingVars)
  xcolumn <- sym(xval)
  ycolumn <- sym(yval)
  # title defaults for saving plot
  label_title = ""
  points_title = ""
  summary_title = ""

  samples_resp_plot <- resp_dat %>%
    select(`SampleID`, contains("Day"), contains("Time"),
           `TotalCO2 (ppm)`:PercentCH4,
           all_of(GroupingVars))

  samples_resp_plot_summary <- samples_resp_plot %>%
    group_by(across(all_of(c(xval, GroupingVars)))) %>%
    select(contains(c("CO2", "CH4"))) %>%
    summarize(across(contains(c("CH4", "CO2")), list(mean = ~mean(., na.rm = T), sd = ~sd(., na.rm = T),
                                             N = ~sum(!is.na(.)), se = ~sd(., na.rm = T)/sqrt(sum(!is.na(.))),
                                             se_lwr = ~mean(., na.rm = T) - (sd(., na.rm = T)/sqrt(sum(!is.na(.)))),
                                             se_upr = ~mean(., na.rm = T) + (sd(., na.rm = T)/sqrt(sum(!is.na(.))))
                                             )
                   )) %>%
    ungroup()

  ymean = sym(paste0(yval, "_mean"))
  ylwr =  sym(paste0(yval, "_se_lwr"))
  yupr =  sym(paste0(yval, "_se_upr"))
  color_var = sym(GroupingVars[2])
  #shape_var = sym(GroupingVars[3])

  if(summary_plot){
    summary_plot <- samples_resp_plot_summary %>%
      ggplot(aes(y = !!ymean, x = !!xcolumn, color = !!color_var)) +
      geom_point(aes_string(shape = "SiteIncRound"), size = 3) +
      scale_shape_manual(values = c(21,16,24,17,3)) +
      geom_line(aes_string(group = GroupingVars[2])) +
      geom_errorbar(data = samples_resp_plot_summary,
                  aes(ymin = !!ylwr,
                       ymax = !!yupr))
  } else {
    summary_plot <- samples_resp_plot_summary %>%
      ggplot(aes(y = !!ymean, x = !!xcolumn, color = !!color_var)) +
      scale_shape_manual(values = c(21,16,24,17,3))
    summary_title = "_noSummary"
  }
  
  # Plot all points as well as means and std.errs
  if(all_points) {
    summary_plot <- summary_plot +
      geom_point(data = samples_resp_plot,
               aes(y = !!ycolumn, x = !!xcolumn, shape = SiteIncRound))
    points_title = "_all_points"
  }
    
  # Plot jar labels
  if(jar_labels) {
    summary_plot <- summary_plot +
    geom_text_repel(data = samples_resp_plot,
               aes(y = !!ycolumn,
                   label = `SampleID`))
    label_title = "_labels"
    }

  if(save) {
    ggsave(filename = paste0(save_location, yval, points_title, label_title, summary_title, ".png"),
           width = 7, height = 5,
           summary_plot,
           device = "png")
  }

  return(summary_plot)

}
```

```{r}

# Prepare grouping variables in data
QC_group_by_site_coal <- samples_resp %>%
  ungroup() %>%
  separate(`SampleID`, into = c("SiteAbbr", "MeasureType", "JarNum"), 
           sep = "-", remove = F) %>%
  rename(Round = `Round Subset (A,B,C,D)`) %>%
  mutate(SiteAbbrPermCoal = paste(SiteAbbr, `Permafrost or Coalescence?`)) %>%
  mutate(SiteIncRound = paste(Site, "- Round", `Incubation Round`)) %>%
  filter(!is.na(SiteAbbrPermCoal))

# Create plots
site_coal_ch4_plot <- plot_resp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CH4_per_g_soil_per_day",
                  xval = "PreIncubationDay",
                  summary_plot = T,
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal"))

site_coal_ch4_plot

site_coal_co2_plot <- plot_resp(resp_dat = QC_group_by_site_coal,
                   yval = "ug_C_CO2_per_g_soil_per_day",
                   xval = "PreIncubationDay",
                   summary_plot = T,
                   jar_labels = F,
                   all_points = T,
                   GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal"))
site_coal_co2_plot

site_coal_percCH4_plot <- plot_resp(resp_dat = QC_group_by_site_coal,
                  yval = "PercentCH4",
                  xval = "PreIncubationDay",
                  summary_plot = T,
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal"))

site_coal_percCH4_plot


site_coal_ch4_plot_label <- plot_resp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CH4_per_g_soil_per_day",
                  xval = "PreIncubationDay",
                  jar_labels = T, 
                  all_points = T,
                  GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal"))
site_coal_ch4_plot_label

site_coal_co2_plot_label <- plot_resp(resp_dat = QC_group_by_site_coal,
                   yval = "ug_C_CO2_per_g_soil_per_day",
                   xval = "PreIncubationDay",
                   jar_labels = T,
                   all_points = T,
                   GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal"))
site_coal_co2_plot_label

site_coal_percCH4_plot_label <- plot_resp(resp_dat = QC_group_by_site_coal,
                   yval = "PercentCH4",
                   xval = "PreIncubationDay",
                   jar_labels = T,
                   all_points = T,
                   GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal"))
site_coal_percCH4_plot_label

site_coal_ch4_plot_pointsonly <- plot_resp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CH4_per_g_soil_per_day",
                  xval = "PreIncubationDay",
                  summary_plot = F,
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal"))

site_coal_ch4_plot_pointsonly

site_coal_co2_plot_pointsonly <- plot_resp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CO2_per_g_soil_per_day",
                  xval = "PreIncubationDay",
                  summary_plot = F,
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal"))
site_coal_co2_plot_pointsonly

site_coal_percCH4_plot_pointsonly <- plot_resp(resp_dat = QC_group_by_site_coal,
                  yval = "PercentCH4",
                  xval = "PreIncubationDay",
                  summary_plot = F,
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("SiteIncRound", "SiteAbbrPermCoal"))
site_coal_percCH4_plot_pointsonly
```

Just for fun, plot each sample with the name of the person who ran it:

```{r}
# By Person
student_data <- QC_group_by_site_coal %>%
  mutate(Names = str_extract(Comments, "[^;]+"),
         Names = gsub("(.+?)([fF]lush.*)", "\\1",Names),
         Names = gsub("Flush times added 02/10 - hhm", NA, Names),
         Names = gsub("Measured by ", "", Names))
site_coal_ch4_plot_pointsonly +
  geom_text_repel(data = student_data, aes(label = Names, y = ug_C_CH4_per_g_soil_per_day))

ggsave(filename = paste0(fp_figout, "ug_C_CH4_per_g_soil_per_day", 
                         "scientist_names", ".png"),
           width = 7, height = 5,
           device = "png")
site_coal_co2_plot_pointsonly +
  geom_text_repel(data = student_data, aes(label = Names, y = ug_C_CO2_per_g_soil_per_day))
ggsave(filename = paste0(fp_figout, "ug_C_CH4_per_g_soil_per_day", 
                         "scientist_names", ".png"),
           width = 7, height = 5,
           device = "png")
```

# How do these data compare to Ernakovich 2017?

```{r}

if(exists("ernakovich_2017")) {
  ernakovich_2017_filt <- ernakovich_2017 %>%
  filter(grepl("anoxic", Treatment)) %>%
  filter(`Day of Incubation` < max(QC_group_by_site_coal$PreIncubationDay, na.rm = T)) %>%
  mutate(`Day of Incubation` = ifelse(grepl("15C", Treatment), 
                                      `Day of Incubation` + 0.5, `Day of Incubation`)) %>% # to make points not overlap in plot
  rename(PreIncubationDay = `Day of Incubation`,
         ug_C_CH4_per_g_soil_per_day_mean = ug_C_CH4_per_g_soil_per_day,
         ug_C_CO2_per_g_soil_per_day_mean = ug_C_CO2_per_g_soil_per_day)
  
# Plotting layers to make ernakovich data consistently plotted on graphs
ernakovich2017_plotting_layers <- list(
    geom_point(data = ernakovich_2017_filt,
               aes(fill = Treatment), shape = 23,color = "NA"),
    scale_fill_manual(name = "Ernakovich et al. 2017", values = c("red", "blue"))
  )


site_coal_ch4_plot +
  geom_errorbar(data = ernakovich_2017_filt %>% filter(Treatment == "anoxic 1C"),
                aes(ymax = ug_C_CH4_per_g_soil_per_day_mean + ug_C_CH4_per_g_soil_per_day_se,
                ymin = ug_C_CH4_per_g_soil_per_day_mean - ug_C_CH4_per_g_soil_per_day_se,
                fill = Treatment), color="blue") +
  geom_errorbar(data = ernakovich_2017_filt %>% filter(Treatment == "anoxic 15C"),
                aes(ymax = ug_C_CH4_per_g_soil_per_day_mean + ug_C_CH4_per_g_soil_per_day_se,
                ymin = ug_C_CH4_per_g_soil_per_day_mean - ug_C_CH4_per_g_soil_per_day_se,
                fill = Treatment), color="red") +
  ernakovich2017_plotting_layers

ggsave(filename = paste0(fp_figout, "ug_C_CH4_per_g_soil_per_day", 
                         "Ernkovich2017_compare", ".png"),
           width = 7, height = 5,
           device = "png")

site_coal_co2_plot +
  geom_errorbar(data = ernakovich_2017_filt %>% filter(Treatment == "anoxic 1C"),
                aes(ymax = ug_C_CO2_per_g_soil_per_day_mean + ug_C_CO2_per_g_soil_per_day_se,
                ymin = ug_C_CO2_per_g_soil_per_day_mean - ug_C_CO2_per_g_soil_per_day_se,
                fill = Treatment), color="blue") +
  geom_errorbar(data = ernakovich_2017_filt %>% filter(Treatment == "anoxic 15C"),
                aes(ymax = ug_C_CO2_per_g_soil_per_day_mean + ug_C_CO2_per_g_soil_per_day_se,
                ymin = ug_C_CO2_per_g_soil_per_day_mean - ug_C_CO2_per_g_soil_per_day_se,
                fill = Treatment), color="red") +
  ernakovich2017_plotting_layers

ggsave(filename = paste0(fp_figout, "ug_C_CO2_per_g_soil_per_day", 
                         "Ernkovich2017_compare", ".png"),
           width = 7, height = 5,
           device = "png")
}
  
```


# More in-Depth Quality Control Practices

### Oxygen treatments must be oxic

Check that no ppm has exceeded 2% CO2 in oxic treatments **CODE NOT CURRENTLY TESTED**

```{r}

samples_resp %>%
  filter(AnoxicOxic == "Oxic") %>%
  filter(!is.na(umol_CO2_per_molAir)) %>%
  filter(IncubationDay >= 0) %>%
  mutate(`Jar has gone anoxic` = ifelse(umol_CO2_per_molAir < 20000, F, T)) %>%
  ggplot(aes(x = IncubationTime/(24*60*60), y = umol_CO2_per_molAir)) +
  geom_point(aes(color = `Jar has gone anoxic`), alpha = 0.5) +
  geom_text_repel(aes(label = SampleID)) +
  geom_hline(yintercept = 20000, color = "#B72F40", linetype = "dashed") +
  annotate("curve", xend = 5, x = 5 + 0.5, y = 40000, yend = 20000, curvature = -0.1,
          arrow = arrow(length = unit(0.03, "npc")), color = "#B72F40") +
  annotate("text", x = 5 + 0.5, y = 40000, label = "2% CO2", hjust = 0, vjust = 0, color = "#B72F40") +
  ylab("umol CO2 per mol air (ppm)") + xlab("Day of Incubation") +
  scale_y_continuous(labels = scales::comma) +
  theme(legend.position = "bottom")

```

### Identify and filter out outlier points and jars

```{r}
plot_resp_TEA_temp <- function(resp_dat = QC_group_by_site_coal,
                      yval = "ug_C_CH4_per_g_soil_per_day",
                      xval = "IncubationDay",
                      summary_plot = T,
                      jar_labels = F,
                      all_points = T,
                      save = T,
                      save_location = fp_figout,
                      GroupingVars = c("TEA", "Incubation Temp",
                                       "SiteAbbrPermCoal")) {
  #args_list <- list(...)
  #print(args_list)
  #print(GroupingVars)
  xcolumn <- sym(xval)
  ycolumn <- sym(yval)
  # title defaults for saving plot
  label_title = ""
  points_title = ""
  summary_title = ""

  samples_resp_plot <- resp_dat %>%
    select(`SampleID`, contains("Day"), contains("Time"),
           `TotalCO2 (ppm)`:PercentCH4,
           all_of(GroupingVars))

  samples_resp_plot_summary <- samples_resp_plot %>%
    group_by(across(all_of(c(xval, GroupingVars)))) %>%
    select(contains(c("CO2", "CH4"))) %>%
    summarize(across(contains(c("CH4", "CO2")), 
                     list(mean = ~mean(., na.rm = T),
                          sd = ~sd(., na.rm = T),
                          N = ~sum(!is.na(.)), 
                          se = ~sd(., na.rm = T)/sqrt(sum(!is.na(.))),
                          se_lwr = ~mean(., na.rm = T) - (sd(., na.rm = T)/sqrt(sum(!is.na(.)))),
                          se_upr = ~mean(., na.rm = T) + (sd(., na.rm = T)/sqrt(sum(!is.na(.))))
                          )
                   )) %>%
    ungroup()

  ymean = sym(paste0(yval, "_mean"))
  ylwr =  sym(paste0(yval, "_se_lwr"))
  yupr =  sym(paste0(yval, "_se_upr"))
  line_var = sym(GroupingVars[3])
  color_var = sym(GroupingVars[2])
  shape_var = sym(GroupingVars[3])

  if(summary_plot){
    summary_plot <- samples_resp_plot_summary %>%
      ggplot(aes(y = !!ymean, x = !!xcolumn, color = !!color_var)) +
      geom_point(aes(shape = !!shape_var), size = 3) +
      scale_shape_manual(values = c(21,16,24,17)) +
      scale_color_manual(breaks = temp_levels, values = temp_colors) +
      geom_line(aes(group = !!line_var)) +
      geom_errorbar(data = samples_resp_plot_summary,
                  aes(ymin = !!ylwr,
                       ymax = !!yupr))
  } else {
    summary_plot <- samples_resp_plot_summary %>%
      ggplot(aes(y = !!ymean, x = !!xcolumn, color = !!color_var)) +
      scale_shape_manual(values = c(21,16,24,17)) +
      scale_color_manual(breaks = temp_levels, values = temp_colors)
    summary_title = "_noSummary"
  }
  
  # Plot all points as well as means and std.errs
  if(all_points) {
    summary_plot <- summary_plot +
      geom_point(data = samples_resp_plot,
               aes(y = !!ycolumn, x = !!xcolumn, shape = !!shape_var)) +
      facet_grid(TEA ~`Incubation Temp`, scales = "free_y")
    points_title = "_all_points"
  }
    
  # Plot jar labels
  if(jar_labels) {
    summary_plot <- summary_plot +
    geom_text_repel(data = samples_resp_plot,
               aes(y = !!ycolumn,
                   label = `SampleID`))
    label_title = "_labels"
    }

  summary_plot <- summary_plot +
    theme(legend.position = "bottom", 
          legend.direction = "horizontal",
          legend.title.position = "top")
  if(save) {
    ggsave(filename = paste0(save_location, yval, "_TEATemp", 
                             points_title, label_title, summary_title, ".png"),
           width = 8, height = 8,
           summary_plot,
           device = "png")
  }

  return(summary_plot)

}
```


```{r}

# Prepare grouping variables in data
QC_group_by_site_coal <- samples_resp %>%
  ungroup() %>%
  filter(IncubationDay >= 0) %>%
  filter(`Incubation Temp` %in% c("4", "10", "15", "20", "25")) %>%
  separate(`SampleID`, into = c("SiteAbbr", "MeasureType", "JarNum"), 
           sep = "-", remove = F) %>%
  rename(Round = `Round Subset (A,B,C,D)`) %>%
  mutate(SiteAbbrPermCoal = paste(SiteAbbr, `Permafrost or Coalescence?`)) %>%
  mutate(SiteIncRound = paste(Site, "- Round", `Incubation Round`)) %>%
  filter(!is.na(SiteAbbrPermCoal))

# Create plots
site_coal_ch4_plot <- plot_resp_TEA_temp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CH4_per_g_soil_per_day",
                  xval = "IncubationDay",
                  summary_plot = T,
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("TEA", "Incubation Temp", 
                                   "SiteAbbrPermCoal"))

site_coal_ch4_plot

site_coal_co2_plot <- plot_resp_TEA_temp(resp_dat = QC_group_by_site_coal,
                   yval = "ug_C_CO2_per_g_soil_per_day",
                   xval = "IncubationDay",
                   summary_plot = T,
                   jar_labels = F,
                   all_points = T,
                   GroupingVars = c("TEA", "Incubation Temp", 
                                   "SiteAbbrPermCoal"))
site_coal_co2_plot

site_coal_percCH4_plot <- plot_resp_TEA_temp(resp_dat = QC_group_by_site_coal,
                  yval = "PercentCH4",
                  xval = "IncubationDay",
                  summary_plot = T,
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("TEA", "Incubation Temp", 
                                   "SiteAbbrPermCoal"))

site_coal_percCH4_plot

site_coal_ch4_plot_label <- plot_resp_TEA_temp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CH4_per_g_soil_per_day",
                  xval = "IncubationDay",
                  jar_labels = T, 
                  all_points = T,
                  GroupingVars = c("TEA", "Incubation Temp", 
                                   "SiteAbbrPermCoal"))
site_coal_ch4_plot_label

site_coal_co2_plot_label <- plot_resp_TEA_temp(resp_dat = QC_group_by_site_coal,
                   yval = "ug_C_CO2_per_g_soil_per_day",
                   xval = "IncubationDay",
                   jar_labels = T,
                   all_points = T,
                  GroupingVars = c("TEA", "Incubation Temp", 
                                   "SiteAbbrPermCoal"))
site_coal_co2_plot_label

site_coal_percCH4_plot_label <- plot_resp_TEA_temp(resp_dat = QC_group_by_site_coal,
                   yval = "PercentCH4",
                   xval = "IncubationDay",
                   jar_labels = T,
                   all_points = T,                  
                   GroupingVars = c("TEA", "Incubation Temp", 
                                   "SiteAbbrPermCoal"))
site_coal_percCH4_plot_label

site_coal_ch4_plot_pointsonly <- plot_resp_TEA_temp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CH4_per_g_soil_per_day",
                  xval = "IncubationDay",
                  summary_plot = F,
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("TEA", "Incubation Temp", 
                                   "SiteAbbrPermCoal"))

site_coal_ch4_plot_pointsonly

site_coal_co2_plot_pointsonly <- plot_resp_TEA_temp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CO2_per_g_soil_per_day",
                  xval = "IncubationDay",
                  summary_plot = F,
                  jar_labels = F, 
                  all_points = T,                  
                  GroupingVars = c("TEA", "Incubation Temp", 
                                   "SiteAbbrPermCoal"))
site_coal_co2_plot_pointsonly

site_coal_percCH4_plot_pointsonly <- plot_resp_TEA_temp(resp_dat = QC_group_by_site_coal,
                  yval = "PercentCH4",
                  xval = "IncubationDay",
                  summary_plot = F,
                  jar_labels = F, 
                  all_points = T,                  
                  GroupingVars = c("TEA", "Incubation Temp", 
                                   "SiteAbbrPermCoal"))
site_coal_percCH4_plot_pointsonly 


QC_group_by_site_coal %>%
  filter_out(Timestamp == "2026-04-30 10:13:00") %>%
#   filter_out(JarIDNum == "225" & Date == "2026-04-30") %>%
      ggplot(aes(y = ug_C_CH4_per_g_soil_per_day, x = IncubationDay, color = `Incubation Temp`)) +
      scale_shape_manual(values = c(21,16,24,17)) +
      scale_color_manual(breaks = temp_levels, values = temp_colors) +
      geom_line(aes(group = JarIDNum)) +
      geom_text_repel(aes(group = JarIDNum, label = JarIDNum)) +
      geom_point(aes(shape = SiteAbbrPermCoal)) +
      facet_grid(TEA ~`Incubation Temp`)

QC_group_by_site_coal %>%
  filter_out(Timestamp == "2026-04-30 10:13:00") %>%
      ggplot(aes(y = ug_C_CO2_per_g_soil_per_day, x = IncubationDay, color = `Incubation Temp`)) +
      scale_shape_manual(values = c(21,16,24,17)) +
      scale_color_manual(breaks = temp_levels, values = temp_colors) +
      geom_line(aes(group = JarIDNum)) +
      geom_point(aes(shape = SiteAbbrPermCoal)) +
      geom_text_repel(aes(group = JarIDNum, label = JarIDNum)) +
      facet_grid(TEA ~`Incubation Temp`, scales = "free_y")
  


```



# Calculate and plot Cumulative C

5.  **NOT CURRENTLY IMPLEMENTED:** Calculate the cumulative sum of carbon released by multiplying the gas flux rate by the time that has passed for each time step, and cumulatively adding all of these values together.

# Convert Licor and picarro readings into same space.

-   will need to filter out and/or average across samples that are measured on both instruments.

# Once QC is finished, write out the final data
