---
title: "Quality Control of Respiration measurements"
format: html
editor: visual
---

```{r}
#| message: false
#| warning: false
#| include: false
knitr::opts_chunk$set(echo = T, eval = T, message = F, warning = F)
# load libraries
library(tidyverse)
library(here)
library(lubridate)
library(ggrepel)
library(broom)

# Set ggplot theme
theme_set(theme_bw())
```

## 

This script provides a guide for how to perform QA/QC on respiration measurements

## User setup

**Instructions for SEI users**: 1) Download the "\*-Respiration.xlsx" spreadsheets to a folder called something like `data/licor/` or `data/picarro/` depending on the instrument that the measurements came from. All respiration spreadsheets should have names that start with a date in `YYYY-MM-DD` format, and end with `_Respiration.xlsx`

2)  Specify the location of spreadsheets and jar metadata as well. The `26SEI_JarLabelingandLoadingSpreadsheet_all_jars.tsv` combines both the Healy and Ice Cut samples and is generated by the utility script "SEI_metadata_reformat.R" (found in the `utilities` folder). To run the utility script you'll need the "26SEI_JarLabelingandLoadingSpreadsheet.xlsx" and "GasMeasurementTracking_SEI26(Flush 1).csv" tables from sharepoint

```{r user-defined-variables}

# location of sample metadata sheet
fp_metadata <- "~/Downloads/26SEI_JarLabelingandLoadingSpreadsheet_all_jars.tsv"

# location of sample respiration spreadsheets
fp_licor <- "~/Downloads/OneDrive_1_2-2-2026/LicorData_PreInc_SEI26/"

# location of sample respiration spreadsheets
fp_picarro <- "~/Downloads/OneDrive_1_2-2-2026/PicarroData_PreInc_SEI26/" # temporary; eventually move to examples
```

```{r useful-constants-functions}

# Define Constants
GasConstant_R <- 0.0821 # L*atm*K^-1*mol^-1
PSI_to_atm <- 0.06804596 # psi to atm

# Outlier function
isnt_out_tukey <- function(x, k = 1.5, na.rm = TRUE) {
  quar <- quantile(x, probs = c(0.25, 0.75), na.rm = na.rm)
  iqr <- diff(quar)
  
  (quar[1] - k * iqr <= x) & (x <= quar[2] + k * iqr)
}
```

### Reading in the sample metadata and jar information

First we'll read in the sample metadata and jar information. This is important because it contains the pre-incubation and incubation start times.

```{r read-sample-data}
# Reading in sample metadata information
jar_info_raw <- read_tsv(file = fp_metadata) %>%
  mutate(across(JarNumber:`Specimen Cup Number`, ~ as.factor(.x))) %>%
  rename(AnoxicOxic =`Anoxic/Oxic?`)  

# Extract incubation start times
jar_incubation_time_info <- jar_info_raw %>%
  select(SampleID, contains("IncubationStart")) %>%
  mutate(IncubationStart = as_datetime("2026-02-09 12:00:00", tz = "America/New_York")) %>% # temporary until start date is determined added in
  rename(`Jar Number` = 1)

# Extract data relevant for QC of samples, plotting, and calculation of rates 
jar_info <- jar_info_raw %>%
  select(SampleID:Site, `Mass of Soil(g Lab Scale)`, GSC_Perc, HeadspaceVolume)


```

Below we create a function for reading in respiration data and doing basic reformatting that allows the picarro and licor respiration measurements to be combined. These include: 1. Converting picarro readings from 12-C and 13-C into total C 2. Making sure all the columns have the right formats in R 3. Ensuring the dates and times are correctly formatted, and in the right time zone 4. Converting licor and picarro information into each other with a scaling factor. *Does not do this currently - but will eventually*

```{r read-resp-data}
read_respiration_files <- function(resp_file_path_licor = NULL,
                                   resp_file_path_picarro = NULL, 
                                   sample_id_column = "Jar Number",
                                   incubation_start=jar_incubation_time_info,
                                   file_pattern = ".xlsx") {
#' Function: read_respiration_files()
#' Takes a folder containing either Picarro or Licor-style respiration templates
#' and reads them in, does baisc cleaning functions on them. 
# @param resp_file_path_licor a directory path where identically formatted respiration spreadsheets for the licor are housed. If you do not have data from the instrument, leave argument blank.
# @param resp_file_path_picarro a directory path where identically formated respiration spreadsheets for the picarro are housed. If you do not have data from the instrument, leave argument blank.
# @param sample_id_column defaults to "Jar Number", change only if you have changed the sample id column name. A character.
# @param incubation_start A dataframe that includes the sample ID column and the time of the incubation start. 
# @return a list of data frames containing the licor data, the picarro data, and a data frame that combines both.
# @examples tbd
  
  # Dependencies
  library(tidyverse)
  library(lubridate)
  library(readxl)
  
  # Constants
  PSI_to_atm <- 0.06804596 # for converting headspace pressure to moles of gas
  
  # Return list
  return_list <- NA
  
  # Testing
  # resp_file_path_licor = "~/Downloads/licor_data"
  # resp_file_path_picarro = "~/Downloads/OneDrive_1_1-20-2026"
  # incubation_start <- "2026-01-09 12:00:00" 
  # file_pattern = ".xlsx"
  # Check arguments for validity and warn user about states  
  stopifnot((!is.null(resp_file_path_licor) | !is.null(resp_file_path_picarro)))
  
  # Initial read-in
  if(!is_null(resp_file_path_licor)) {
    # print("here") # Testing
    Resp_files <- list.files(here(resp_file_path_licor),
                             pattern = file_pattern, full.names = T)
    
    col_types_licor <- c(rep("text", 2), #SampleType:JarNumber
                         rep("numeric", 5), #RoomTemperature:ConcentrationOfStandard
                         "date", # Time last flushed
                         rep("text", 3), # Comments, id, si.no
                         "date", # timestamp
                         rep("numeric", 4), # mean-pre:Est Conc
                         "text", # Remarks
                         rep("numeric", 10), #:effective volume of loop:True amount of CO2
                       "guess","guess","guess","guess") # standard curve calcs
    
    resp_raw_licor <- map_dfr(Resp_files, 
                               ~readxl::read_excel(.x,sheet = 1,
                                                   col_types = col_types_licor,
                                                   na = c("nan", "#VALUE!", "#DIV/0!"))) %>% 
      filter(!is.na(SampleType)) %>%
      select(SampleType:`True amount of CO2 (uMol)`) %>%
      mutate(MeasurementInstrument = "Licor",
             Timestamp = as_datetime(Timestamp, tz = "America/New_York")) # need to check that this is correct!
    
    # Change names to match variables between licor and picarro
    resp_raw_licor <- resp_raw_licor %>%
      mutate(`TotalCO2 (ppm)` = `True amount of CO2 (uMol)`/ `Moles of Air in sample`) %>%
      rename(`Concentration of Standard Curve CO2 Gas (ppm)` = `Concentration of Standard Curve Gas (ppm)`,
             `TotalCO2 (uMol)` = `True amount of CO2 (uMol)`) %>%
      mutate(JarPressure_atm = `Jar pressure before sampling (psi)`*PSI_to_atm) %>%
      # calcualte durations:
      left_join(incubation_start, by = c("Jar Number")) %>%
      mutate(Date = as_date(Timestamp),
             IncubationDay = Date - as_date(IncubationStart),
             IncubationTime = Timestamp - as_datetime(IncubationStart, tz = "America/New_York"),
             PreIncubationDay = Date - as_date(PreIncubationStart),
             PreIncubationTime = Timestamp - as_datetime(PreIncubationStart, tz = "America/New_York"),
             TimeSinceZerod = Timestamp - `Time Incubation Jar was Last Flushed (Leave blank for standard curve)`)
  }
  
  if(exists("resp_file_path_picarro")) {
    Resp_files <- list.files(here(resp_file_path_picarro),
                             pattern = file_pattern, full.names = T)
    
    col_types_picarro <- c(rep("text", 2), #sampletype:jar number
                       rep("numeric", 16), #RoomTemperature:EffectiveIsotopicRatio
                       "date", #date last flushed
                       "date", #time last flushed
                       rep("text", 1), # comments
                       "date", # Date
                       "date", # time of start of measurements
                       rep("numeric", 8), #Mean of pre-12CO2:MeanH2O%post 
                       "text", # Remarks 
                       rep("numeric", 32), # effective volume of loop:TrueDelta permil
                       rep("guess", 12)) # regression calculations
    # readxl::read_excel(Resp_files[2], sheet = 1, skip = 1,
    #                                                col_types = col_types_picarro,
    #                                                na = c("nan", "#VALUE!", "#DIV/0!", "#REF!"))
    
    # Read in data
    resp_raw_picarro <- map_dfr(Resp_files, 
                               ~readxl::read_excel(., sheet = 1, skip = 1,
                                                   col_types = col_types_picarro,
                                                   range = cell_limits(c(2,1), c(NA,76)),
                                                   na = c("nan", "#VALUE!", "#DIV/0!", "#REF!"))) %>% 
      filter(!is.na(SampleType)) %>%
      select(SampleType:`True delta iCO2 (permil)`) %>%
      mutate(MeasurementInstrument = "picarro")
    
    # Fix timestamps for measurement and flush time
    date(resp_raw_picarro$`Time of start of measurement`) <- resp_raw_picarro$Date 
    resp_raw_picarro$`Time of start of measurement` <- force_tz(resp_raw_picarro$`Time of start of measurement`, "America/New_York") 
    
    date(resp_raw_picarro$`Time Incubation Jar was Last Flushed (Leave blank for standard curve)`) <- resp_raw_picarro$`Date Incubation Jar was Last Flushed (Leave blank for standard curve)` 
    resp_raw_picarro$`Time Incubation Jar was Last Flushed (Leave blank for standard curve)` <- force_tz(resp_raw_picarro$`Time Incubation Jar was Last Flushed (Leave blank for standard curve)`, "America/New_York") 
    
    # Create columns for total CH4 and CO2
    resp_raw_picarro <- resp_raw_picarro %>% 
      mutate(`atom%12CO2 based on delta` = 1 - `Uncalibrated apparent atom% 13CO2 based on delta value`,
             `atom%12CH4 based on delta` = 1 - `Uncalibrated apparent atom% 13CH4 based on delta value`,
             `TotalCO2 (ppm)` = `Uncalibrated apparent 12CO2 concentration (ppm)` / `atom%12CO2 based on delta`,
             `TotalCH4 (ppm)` = `Uncalibrated apparent 12CH4 concentration (ppm)` / `atom%12CH4 based on delta`,
             `TotalCO2 (uMol)` = `TotalCO2 (ppm)` * `Moles of Air in sample`,
             `TotalCH4 (uMol)` = `TotalCH4 (ppm)` * `Moles of Air in sample`) %>%
      # Create one timestamp
      rename(Timestamp = `Time of start of measurement`) %>%
      # convert jar pressure to atm:
      mutate(JarPressure_atm = `Jar pressure before sampling (psi)`*PSI_to_atm) %>%
      # calculate durations
      left_join(incubation_start, by = c("Jar Number")) %>%
      mutate(Date = as_date(Timestamp),
             IncubationDay = Date - as_date(IncubationStart),
             IncubationTime = Timestamp - as_datetime(IncubationStart, tz = "America/New_York"),
             PreIncubationDay = Date - as_date(PreIncubationStart),
             PreIncubationTime = Timestamp - as_datetime(PreIncubationStart, tz = "America/New_York"),
             TimeSinceZerod = Timestamp - `Time Incubation Jar was Last Flushed (Leave blank for standard curve)`)
  }
  
  
  # Next steps:
  # Determine which of the resp_raw data is present (use exists(resp_raw_picarro))
  # combine data together into one dataframe
  # Create the incubation day and time columns, use either the user-specified beginning datetime or the minimum date
  
  if(exists("resp_raw_licor") & exists("resp_raw_picarro")) {
    resp_raw <- bind_rows(resp_raw_licor, resp_raw_picarro)
    return_list <- list(licor_resp_raw = resp_raw_licor,
                        picarro_resp_raw = resp_raw_picarro)
  } else if(exists("resp_raw_licor")) {
    resp_raw <- resp_raw_licor
    return_list <- list(licor_resp_raw = resp_raw_licor,
                    picarro_resp_raw = NA)
  } else if(exists("resp_raw_picarro")) {
      resp_raw <- resp_raw_picarro
      return_list <- list(licor_resp_raw = NA,
                    picarro_resp_raw = resp_raw_picarro)
  }
  
  
  # Calculate durations and jar pressures
    resp_raw <- resp_raw %>%  
      mutate(SampleType = factor(SampleType),
             `Jar Number` = factor(as.character(`Jar Number`)))
  
  # Return statement
  return_list$resp_raw <- resp_raw
  return(return_list)  
}

```

Run the function to create the input.

```{r}
resp_dat_raw <- read_respiration_files(
  resp_file_path_licor = fp_licor,
  resp_file_path_picarro = fp_picarro,
  incubation_start = jar_incubation_time_info,
  file_pattern = ".xlsx")

```

### Calculate Respirations

Next we need to calculate the respiration rates. In brief this means we need to:

1.  Correct the amount of gas measured for the volume of air in the headspace (which varies between jars)
2.  Calculate the dry soil weight in each jar (based on the gravimetric water content or gravimetric soil content (GSC_perc) )
3.  Convert the total gas produced in micromoles to micrograms (to align with conventional units)
4.  Divide the amount of gas produced by the hours and days since the jars have been flushed to produce and hourly and daily rate


```{r, respiration-calculations}

samples_resp <- resp_dat_raw$picarro_resp_raw %>%
  filter(SampleType == "Sample") %>%
  # Convert into a per-gram basis, based on Jar number
  left_join(jar_info, by = c("Jar Number" = "SampleID")) %>%
  # If jar pressure is 0 PSI make it a very small number rather than zero:
  mutate(`Jar pressure before sampling (psi)` = ifelse(`Jar pressure before sampling (psi)` == 0, 0.001, `Jar pressure before sampling (psi)`)) %>%
  mutate(jar_headspace_mL = HeadspaceVolume, # need to correct for specimen cup volume
         soil_weight_dry = `Mass of Soil(g Lab Scale)` * (GSC_Perc/100),
         MolAir_headspace = (1/1000)*(jar_headspace_mL * JarPressure_atm)/(GasConstant_R * `Room Temperature (K)`),
         umol_CO2_in_headspace = jar_headspace_mL * (`TotalCO2 (uMol)`)/`Volume of sample or standard (mL of gas measured)`,
         umol_CH4_in_headspace = jar_headspace_mL * (`TotalCH4 (uMol)`)/`Volume of sample or standard (mL of gas measured)`,
         umol_CO2_per_molAir = umol_CO2_in_headspace/MolAir_headspace,
         umol_CH4_per_molAir = umol_CH4_in_headspace/MolAir_headspace,
         ug_C_CO2_in_headspace = jar_headspace_mL * (`TotalCO2 (uMol)` * 12.01)/`Volume of sample or standard (mL of gas measured)`,
         ug_C_CH4_in_headspace = jar_headspace_mL * (`TotalCH4 (uMol)` * 12.01)/`Volume of sample or standard (mL of gas measured)`,
         ug_C_CO2_per_g_soil = ug_C_CO2_in_headspace/soil_weight_dry,
         ug_C_CH4_per_g_soil = ug_C_CH4_in_headspace/soil_weight_dry) %>%
  # make it a rate just using temporary numbers here
  mutate(ug_C_CO2_per_g_soil_per_hour = ug_C_CO2_per_g_soil/as.numeric(TimeSinceZerod, unit = "hours"),
         ug_C_CO2_per_g_soil_per_day = ug_C_CO2_per_g_soil/(as.numeric(TimeSinceZerod, unit = "days"))) %>%
  mutate(ug_C_CH4_per_g_soil_per_hour = ug_C_CH4_per_g_soil/as.numeric(TimeSinceZerod, unit = "hours"),
         ug_C_CH4_per_g_soil_per_day = ug_C_CH4_per_g_soil/(as.numeric(TimeSinceZerod, unit = "days")))


```

# Basic checks:

Any samples NA? figure out why before moving on:
```{r}
rate_cols <- c("ug_C_CO2_per_g_soil_per_hour", "ug_C_CO2_per_g_soil_per_day",
         "ug_C_CH4_per_g_soil_per_hour","ug_C_CH4_per_g_soil_per_day") 


samples_resp %>%
  filter(if_any(all_of(rate_cols), ~is.na(.x))) %>% View()

```

Common issues:

- Flush date and time are not entered correctly
- temperature, pressure, or other necessary variable is missing.
- soil weights not entered
- Typo in sample name or number

If necessary, find the date of the timestamp for the measurement and check the corresponding excel sheet and lab notebook pages. Message the person who took the measurement to verify any uncertainties. 

**Only make changes to the raw spreadsheets if you are certain you know the correct value. And always note your changes in a comment column, with your name or initials and the date.**

# Plot respiration

```{r}

plot_resp <- function(resp_dat = samples_resp,
                      yval = "ug_C_CH4_per_g_soil",
                      xval = "PreIncubationDay",
                      jar_labels = F,
                      all_points = T,
                      save = T,
                      GroupingVars = c("Site", "SiteAbbrPermCoal")) {
  #args_list <- list(...)
  #print(args_list)
  #print(GroupingVars)
  xcolumn <- sym(xval)
  ycolumn <- sym(yval)
  label_title = ""
  points_title = ""

  samples_resp_plot <- resp_dat %>%
    select(`Jar Number`, contains("Day"), contains("Time"),
           `TotalCO2 (ppm)`:ug_C_CH4_per_g_soil_per_day,
           all_of(GroupingVars))

  samples_resp_plot_summary <- samples_resp_plot %>%
    group_by(across(all_of(c(xval, GroupingVars)))) %>%
    select(contains(c("CO2", "CH4"))) %>%
    summarize(across(contains(c("CH4", "CO2")), list(mean = ~mean(., na.rm = T), sd = ~sd(., na.rm = T),
                                             N = ~sum(!is.na(.)), se = ~sd(., na.rm = T)/sqrt(sum(!is.na(.))),
                                             se_lwr = ~mean(., na.rm = T) - (sd(., na.rm = T)/sqrt(sum(!is.na(.)))),
                                             se_upr = ~mean(., na.rm = T) + (sd(., na.rm = T)/sqrt(sum(!is.na(.))))
                                             )
                   )) %>%
    ungroup()

  ymean = sym(paste0(yval, "_mean"))
  ylwr =  sym(paste0(yval, "_se_lwr"))
  yupr =  sym(paste0(yval, "_se_upr"))
  color_var = sym(GroupingVars[2])

  summary_plot <- samples_resp_plot_summary %>%
    ggplot(aes(y = !!ymean, x = !!xcolumn, color = !!color_var)) +
    geom_point(aes_string(shape = "Site"), size = 3) +
    geom_line(aes_string(group = GroupingVars[2])) +
    geom_errorbar(data = samples_resp_plot_summary,
                  aes(ymin = !!ylwr,
                       ymax = !!yupr))
  
  # Plot all points as well as means and std.errs
  if(all_points) {
    summary_plot <- summary_plot +
      geom_point(data = samples_resp_plot,
               aes(y = !!ycolumn, x = !!xcolumn))
    points_title = "_all_points"
  }
    
  # Plot jar labels
  if(jar_labels) {
    summary_plot <- summary_plot +
    geom_text_repel(data = samples_resp_plot,
               aes(y = !!ycolumn,
                   label = `Jar Number`))
    label_title = "_labels"
    }

  if(save) {
    ggsave(filename = paste0("~/Downloads/", yval, points_title, label_title, ".png"),
           width = 7, height = 5,
           summary_plot,
           device = "png")
  }

  return(summary_plot)

}

QC_group_by_site_coal <- samples_resp %>%
  ungroup() %>%
  separate(`Jar Number`, into = c("SiteAbbr", "MeasureType", "JarNum"), 
           sep = "-", remove = F) %>%
  mutate(SiteAbbrPermCoal = paste(SiteAbbr, `Permafrost or Coalescence?`)) %>%
  filter(!is.na(SiteAbbrPermCoal))

site_coal_ch4_plot <- plot_resp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CH4_per_g_soil_per_day",
                  xval = "PreIncubationDay",
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("Site", "SiteAbbrPermCoal"))

site_coal_co2_plot <- plot_resp(resp_dat = QC_group_by_site_coal,
                   yval = "ug_C_CO2_per_g_soil_per_day",
                   xval = "PreIncubationDay",
                   jar_labels = F,
                   GroupingVars = c("Site", "SiteAbbrPermCoal"))

site_coal_ch4_plot_label <- plot_resp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CH4_per_g_soil_per_day",
                  xval = "PreIncubationDay",
                  jar_labels = T, 
                  all_points = T,
                  GroupingVars = c("Site", "SiteAbbrPermCoal"))

site_coal_co2_plot_label <- plot_resp(resp_dat = QC_group_by_site_coal,
                   yval = "ug_C_CO2_per_g_soil_per_day",
                   xval = "PreIncubationDay",
                   jar_labels = T,
                   GroupingVars = c("Site", "SiteAbbrPermCoal"))

```



# Quality control Practices

### Oxygen treatments must be oxic

Check that no ppm has exceeded 2% CO2 in oxic treatments **CODE NOT CURRENTLY TESTED**

```{r}

samples_resp %>%
  filter(AnoxicOxic == "Oxic") %>%
  filter(!is.na(umol_CO2_per_molAir)) %>%
  mutate(Oxic_Conditions = ifelse(umol_CO2_per_molAir < 20000, F, T)) %>%
  ggplot(aes(x = IncubationTime, y = umol_CO2_per_molAir)) +
  geom_point(aes(color = Oxic_Conditions)) +
  geom_text_repel(aes(label = `Jar Number`)) +
  geom_hline(yintercept = 20000, color = "#B72F40", linetype = "dashed") +
  annotate("curve", xend = 5, x = 5 + 0.5, y = 40000, yend = 20000, curvature = -0.1,
          arrow = arrow(length = unit(0.03, "npc")), color = "#B72F40") +
  annotate("text", x = 5 + 0.5, y = 40000, label = "2% CO2", hjust = 0, vjust = 0, color = "#B72F40") +
  ylab("umol CO2 per mol air (ppm)") + xlab("Day of Incubation") +
  scale_y_continuous(labels = scales::comma) +
  theme(legend.position = "bottom")

```

### Identify and filter out outlier points and jars


# Calculate and plot Cumulative C
5.  **NOT CURRENTLY IMPLEMENTED:** Calculate the cumulative sum of carbon released by multiplying the gas flux rate by the time that has passed for each time step, and cumulatively adding all of these values together.

```{r}
# Calculate cumulative rates


# NOTE USE ONLY ON SAMPLES THAT HAVE BEEN CONTINUOUSLY MEASURED!
samples_resp_cumulative <- samples_resp %>%
  select(`Jar Number`, `Time Incubation Jar was Last Flushed (Leave blank for standard curve)`, 
         TimeSinceZerod, Date:soil_weight_dry, ug_C_CO2_per_g_soil,ug_C_CH4_per_g_soil, 
         all_of(rate_cols)) %>%
  group_by(`Jar Number`, JarIDNum, Date) %>%
  # Averaging across TimeSinceZerod fixes days where double measurements are taken on the same jar
  summarise(across(all_of(c(rate_cols, "TimeSinceZerod")), ~mean(., na.rm = T))) %>% 
  arrange(`Jar Number`, Date) %>%
  # Multiply rate by time since flush; timesincezero'd is in decimal days
  mutate(CH4_produced_since_zeroed = ug_C_CH4_per_g_soil_per_hour*(as.numeric(TimeSinceZerod, unit = "hours")),
         CO2_produced_since_zeroed = ug_C_CO2_per_g_soil_per_hour*(as.numeric(TimeSinceZerod, unit = "hours")),
         cum_CH4 = cumsum(ug_C_CH4_per_g_soil_per_hour*(as.numeric(TimeSinceZerod, unit = "hours"))),
         cum_CO2 = cumsum(ug_C_CO2_per_g_soil_per_hour*(as.numeric(TimeSinceZerod, unit = "hours"))),
         cum_CH4_day = cumsum(ug_C_CH4_per_g_soil_per_day*(as.numeric(TimeSinceZerod, unit = "days"))),
         cum_CO2_day = cumsum(ug_C_CO2_per_g_soil_per_day*(as.numeric(TimeSinceZerod, unit = "days"))),
         cum_carbon = cum_CH4 + cum_CO2,
         cum_carbon_day = cum_CH4_day + cum_CO2_day) %>%
  left_join(jar_info, by = c("Jar Number" = "SampleID", "JarIDNum")) %>% 
  left_join(jar_incubation_time_info, by = c("Jar Number")) %>%
  ungroup() %>%
  mutate(IncubationDay = as_date(Date) - as_date(IncubationStart),
         PreIncubationDay = as_date(Date) - as_date(PreIncubationStart))
```

# Plot Cumulative C
```{r}

plot_resp_cumulative <- function(resp_dat = QC_group_by_site_coal_cum,
                      yval = "cum_CH4_day",
                      xval = "PreIncubationDay",
                      jar_labels = F,
                      all_points = T,
                      save = T,
                      GroupingVars = c("Site", "SiteAbbrPermCoal")) {
  #args_list <- list(...)
  #print(args_list)
  #print(GroupingVars)
  xcolumn <- sym(xval)
  ycolumn <- sym(yval)
  label_title = ""
  points_title = ""

  samples_resp_plot <- resp_dat %>%
    select(`Jar Number`, contains("Day"), contains("Time"),
           `CH4_produced_since_zeroed`:cum_carbon_day,
           all_of(GroupingVars))

  samples_resp_plot_summary <- samples_resp_plot %>%
    group_by(across(all_of(c(xval, GroupingVars)))) %>%
    select(contains(c("CO2", "CH4"))) %>%
    summarize(across(contains(c("CH4", "CO2")), list(mean = ~mean(., na.rm = T), sd = ~sd(., na.rm = T),
                                             N = ~sum(!is.na(.)), se = ~sd(., na.rm = T)/sqrt(sum(!is.na(.))),
                                             se_lwr = ~mean(., na.rm = T) - (sd(., na.rm = T)/sqrt(sum(!is.na(.)))),
                                             se_upr = ~mean(., na.rm = T) + (sd(., na.rm = T)/sqrt(sum(!is.na(.))))
                                             )
                   )) %>%
    ungroup()

  ymean = sym(paste0(yval, "_mean"))
  ylwr =  sym(paste0(yval, "_se_lwr"))
  yupr =  sym(paste0(yval, "_se_upr"))
  color_var = sym(GroupingVars[2])

  summary_plot <- samples_resp_plot_summary %>%
    ggplot(aes(y = !!ymean, x = !!xcolumn, color = !!color_var)) +
    geom_point(aes_string(shape = "Site"), size = 3) #+
    # geom_line(aes_string(group = GroupingVars[2])) +
    # geom_errorbar(data = samples_resp_plot_summary,
    #               aes(ymin = !!ylwr,
    #                    ymax = !!yupr))
  
  # Plot all points as well as means and std.errs
  if(all_points) {
    summary_plot <- summary_plot +
      geom_point(data = samples_resp_plot,
               aes(y = !!ycolumn, x = !!xcolumn))
    points_title = "_all_points"
  }
    
  # Plot jar labels
  if(jar_labels) {
    summary_plot <- summary_plot +
    geom_text_repel(data = samples_resp_plot,
               aes(y = !!ycolumn,
                   label = `Jar Number`))
    label_title = "_labels"
    }

  if(save) {
    ggsave(filename = paste0("~/Downloads/", yval, points_title, label_title, "cumul.png"),
           width = 7, height = 5,
           summary_plot,
           device = "png")
  }

  return(summary_plot)

}

QC_group_by_site_coal_cum <- samples_resp_cumulative %>%
  ungroup() %>%
  separate(`Jar Number`, into = c("SiteAbbr", "MeasureType", "JarNum"), 
           sep = "-", remove = F) %>%
  mutate(SiteAbbrPermCoal = paste(SiteAbbr, `Permafrost or Coalescence?`)) %>%
  filter(!is.na(SiteAbbrPermCoal))

site_coal_ch4_cum_plot <- plot_resp_cumulative(resp_dat = QC_group_by_site_coal_cum,
                  yval = "cum_CH4_day",
                  xval = "PreIncubationDay",
                  jar_labels = F, 
                  all_points = T,
                  GroupingVars = c("Site", "SiteAbbrPermCoal"))

site_coal_co2_plot <- plot_resp(resp_dat = QC_group_by_site_coal,
                   yval = "ug_C_CO2_per_g_soil_per_day",
                   xval = "PreIncubationDay",
                   jar_labels = F,
                   GroupingVars = c("Site", "SiteAbbrPermCoal"))

site_coal_ch4_plot_label <- plot_resp(resp_dat = QC_group_by_site_coal,
                  yval = "ug_C_CH4_per_g_soil_per_day",
                  xval = "PreIncubationDay",
                  jar_labels = T, 
                  all_points = T,
                  GroupingVars = c("Site", "SiteAbbrPermCoal"))

site_coal_co2_plot_label <- plot_resp(resp_dat = QC_group_by_site_coal,
                   yval = "ug_C_CO2_per_g_soil_per_day",
                   xval = "PreIncubationDay",
                   jar_labels = T,
                   GroupingVars = c("Site", "SiteAbbrPermCoal"))

```


